---
title: "Simulate summer flounder abundance"
author: "Angelia Miller"
date: "2023-07-25"
output: html_document
---

## Objective ####
Script will simulate summer flounder population and abundances based on stock assessment data

Output: Initial numbers at age, numbers at length, total mortality at age for a set number of years and ages, and recruitment for a given number of years 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)

```

_Stock Assessment Data_
Recruitment at age 0 and fishing mortality data are pulled from the Summer Flounder Management Track Assessment for 2021 (Table 1 and Figure 2). The natural mortality value and Von Bertalanffy growth parameters are pulled from the 66th Stock Assessment Workshop report (2019).
```{r stock data}
set.seed(380)

# Recruitment at age 0 for the most recent 5 years, 2015-2019; numbers were reported in thousands in the MTA report.
Rec_age0 <- c(28416, 33088, 44582, 60598, 48689)*1000

#mean(Rec_age0)
#sd(Rec_age0)

# fishing mortality at MSY (FMSY=F35%) - the target fishing mortality 
# Table 2 - the updated 2021 MTA threshold fishing mortality reference point proxy.
F <- 0.422

# natural mortality
M <- 0.25

#total mortality
Z <- F + M
#mean(Z)
#sd(Z)
#log(sd(Z))

# Von Bertalanffy growth parameters for both male and female
Linf = 83.6
K = 0.14
```


## Simulate Abundance 
```{r sim abundance}

pop <- sim_abundance(ages = 0:7, years = 1:5,
                     R = sim_R(log_mean = log(Rec_age0), log_sd = sd(log(Rec_age0)), plot = TRUE),
                     Z = sim_Z(log_mean = log(Z), log_sd = 0.001, plot = TRUE),
                     growth = sim_vonB(Linf = Linf, K = K, plot = TRUE), 
                     N0 = sim_N0(N0 = "exp", plot = TRUE))
Nage <- tibble(age = 0:7, pop$N0)
plot_surface(pop, mat = "N")

```


```{r, include = FALSE}
### save the data
saveRDS(pop, here("R", "sumflounder", "data", "sumflounder_abundance.rds"))
saveRDS(Nage, here("R", "sumflounder", "data", "sumflounder_Nage.rds"))
```

