---
title: "Performance measure: change in relative abundance over time"
author: "Angelia Miller"
date: "2023-07-25"
output: html_document
---

## Objective 
Script will compare the change in abundance indices over time between the assumed true abundance indices and the estimated abundance indices calculated from `04a-PM-abund-index.Rmd`. 

Output: linear regression slope estimates for the true and estimated abundance indices. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)
```


### Load Data 
```{r}
# read in survey data 
surv_dat1 <- readRDS(here("R", "sumflounder", "data", "sumflounder_survdat.rds"))

obs_stratmu <- readRDS(here("data", "rds", "num-strat-mu_included.rds")) |> filter(SVSPP == 103, EST_YEAR %in% c(2019:2021), SEASON == "SPRING") |> 
  rename(year = EST_YEAR)

# read in the area weights for each strata 
strata_wts <- readRDS(here(sseep.analysis, "data", "rds", "active_strata_wts.rds")) |>
  rename(strat = STRATUM)

# find the total survey area 
BTSArea <- as.integer(sum(strata_wts$Area_SqNm))
```


## True Abundance 
```{r}
obsN_mod <- lm(stratmu~year, data = obs_stratmu) |>
  broom::tidy(conf.int = TRUE) |>
  mutate(TYPE = "Observed") |>
  filter(term == "year")  
  

obsN_mod

```

## Estimated Abundance 
```{r}
estN_mod <- estN_strat |>
  group_by(sim) |>
  nest() |> 
  mutate(model = map(data, ~lm(stratmu~year, data =.)), 
         coef = map(model, broom::tidy, conf.int = TRUE)) |>
  unnest(coef) |> 
  mutate(TYPE = str_c("Simulation", sim, sep = " ")) |>
  filter(term == "year") |>
  select(!c(data, model))

estN_mod

```

## Comparison 
```{r}
obsN_slopes <- bind_rows(obsN_mod, obsN_mod)|> 
  rename(obs_est = estimate, 
         obs_se = std.error, 
         obs_stat = statistic, 
         obs_pval = p.value,
         obs_lcl = conf.low, 
         obs_ucl = conf.high) |> 
  select(!c(term, TYPE))

estN_mod <- estN_mod |> 
  rename(sim_est = estimate, 
         sim_se = std.error, 
         sim_stat = statistic, 
         sim_pval = p.value,
         sim_lcl = conf.low, 
         sim_ucl = conf.high) |> 
  select(!TYPE)

comp <- bind_cols(estN_mod, obsN_slopes)
comp$error <- comp$sim_est - comp$obs_est

```


Slope Comparison
```{r}
Nslopes <- bind_rows(obsN_mod, estN_mod)

Nslopes

ggplot(Nslopes) + 
  geom_pointrange(aes(x = as.factor(TYPE), y = estimate, color = TYPE, ymin=conf.low, ymax = conf.high), position = position_dodge2(width=0.4)) +
  #facet_wrap(vars(SEASON), scales = "free_y") +
  #facet_grid(rows = vars(GEO_AREA), cols = vars(SEASON), scales = "free_y") + 
  labs(x = "Type", y = "Linear Regression Slope Estimates ", color = " ", 
       title = "Changes in Abundance Indices over time") +
  #ylim(0,NA) +
  theme_bw() 
```

## Precision
 - "the precision of the mean density estimate of species g
at time t across replicate surveys and is the standard deviation of the simulated survey estimates (where rgt is the mean density estimate of species g at time t averaged across the D surveys) relative to ygt , the true mean density of species g at time t" (Oyafuso et al 2021)
 - "The True CV (TrueCVst) is the variability of the estimated abundance index across the survey replicates and is defined as the standard deviation of the estimated indices of abundance normalized by the true value" (Oyafuso et al 2023)
```{r}
# trueN_slope <- mean(trueN_mod$estimate)
# 
# estN_mod |>
#   group_by(sim) |> 
#   summarise(cv = sqrt(std.error) / trueN_stratmu)

```

## Bias
```{r}
# diff <- estN_mod |> 
#   group_by(sim) |> 
#   summarise(diff = estimate - trueN_slope)
# 
# # diff |> 
#   group_by(sim) |>
#   summarise(bias = sum(diff)/length(diff))

```


```{r save data, include = FALSE}
# saveRDS(trueN_mod, here("R", "sumflounder", "data", "sumflounder_trueN-mod.rds"))
# 
# saveRDS(estN_mod, here("R", "sumflounder", "data", "sumflounder_estN-mod.rds"))
# 
# saveRDS(Nslopes, here("R", "sumflounder", "data", "sumflounder_slope-changes.rds"))

```

