---
title: "Performance measure: mean length"
author: "Angelia Miller"
date: "2023-07-25"
output: html_document
---

## Objective 
Script will calculate the bias and precision around the stratified mean length measured/collected by the BTS survey.

Output: bias and precision of mean length, adequacy of coverage of length distribution

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)
```

```{r load data, include = FALSE}
# read in survey data 
surv_dat <- readRDS(here("R", "sumflounder", "data", "sumflounder_survdat.rds"))

# read in the area weights for each strata 
strata_wts <- readRDS(here(sseep.analysis, "data", "rds", "active_strata_wts.rds")) |>
  rename(strat = STRATUM)

# find the total survey area 
BTSArea <- as.integer(sum(strata_wts$Area_SqNm))
```


```{r data wrangle, include = FALSE}
#extract data and merge together to obtain lengths in taken in a given strata
setdet <- surv_dat$setdet |> as_tibble()
samp <- surv_dat$samp |> as_tibble()
data <- full_join(setdet, samp, by = "set") 
```

## Calculate the stratified mean length
```{r}
estL_ind <- data |>
    group_by(strat, year, sim) |>
    dplyr::summarise(towct = length(set), # calculate unique tows
              mu = sum(length, na.rm = TRUE)/towct, # find the average biomass based on unique tows rather than observations to avoid potential duplication
              var = ifelse(towct == 1, 0, # if the tow count equals 1, then variance about the mean should be 0
                           sum((length - mu)^2, na.rm = TRUE)/(towct - 1)), 
              .groups = "keep") # if tow count does not equal 1, then find the variance of biomass

# caluclate the weighted average by multiplying by each stratas proportional area
estL_wtd <- estL_ind |>
    left_join(strata_wts, by = "strat") |> # add each stratum area and relative weight to the dataset based on STRATUM number
      group_by(strat, year, sim) |>
    mutate(wt_mu = Area_SqNm * mu, # part one of the stratified mean formula
           wt_var = ((((RelWt)^2) * var) / towct) * (1 - (towct / Area_SqNm)), 
           wt_var = ifelse(is.na(wt_var), 0, wt_var)) 
           # part one of the stratified variance formula

# calculated the stratified average and the standard deviations around the average 
estL_strat <- estL_wtd  |>
    group_by(year, sim) |>
    summarise(stratmu = (sum(wt_mu)) / BTSArea, # part two of the stratified mean formula
              stratvar = sum(wt_var), .groups = "keep") |>
  mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), #logistic standard deviation
         lower = qlnorm(0.025, log(stratmu), sdlog), # lower quantile of the logistic normal distribution
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> # upper quantile of the logistic normal distribution
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), # if sdlog is NaN, replace with 0
         lower = ifelse(is.nan(lower), 0, lower), # if the lower quantile is NaN, replace with 0
         upper = ifelse(is.nan(upper), 0, upper))

estL_strat

```

Plot of the stratified mean length over time
```{r}
ggplot() +
  geom_pointrange(data = estL_strat, aes(x = as.factor(year), y = stratmu, color = as.factor(sim), ymin=lower, ymax = upper), position = position_dodge2(width=0.4)) +
  labs(x = "Year", y = "Stratified Length", color = "Simulation", 
       title = "Stratified Mean Length") +
  #ylim(0,NA) +
  theme_bw() 
```


## Average sampling error for mean length
```{r}
estL_n <- length(samp$length)

estL_se <- estL_strat |> 
  group_by(year, sim) |> 
  summarise(stratsd = sqrt(stratvar), 
            samp_err = 1.96*(stratsd / sqrt(estL_n)))

estL_se |> 
  group_by(sim) |>
  summarise(avg_se = mean(samp_err))

```


## Adequacy of coverage of length distribution
```{r}
ggplot() +
  geom_histogram(data = samp, aes(x = length)) +
  theme_bw() 
```


```{r save data, include = FALSE}
# saveRDS(estL_strat, here("R", "sumflounder", "data", "sumflounder_estL-strat.rds"))

```
