---
title: "Simulate multispecies"
author: "Angelia Miller"
date: "2023-07-25"
output: html_document
---

## Objective ####


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)
library(sf)

sseep.analysis.dat <- "C:/Users/amiller7/Documents/cinar-osse/sseep-analysis/data"

```

Data
```{r}
strata <- readRDS(here(sseep.analysis.dat, "rds", "active_strata_utm.rds"))
wind_areas <- readRDS(here(sseep.analysis.dat, "rds", "wind_areas_062022", "all_wind_areas_Jun2022.rds"))
wind_areas_utm <- st_transform(wind_areas, crs = 32618)

sumflounder <- readRDS(here("R", "sumflounder", "data", "sumflounder_fall-dist.rds"))
atlmackerel <- readRDS(here("R", "atlmackerel", "data", "atlmackerel_dist.rds"))

#scup survey data 
scup_surv_sq <- readRDS(here("data", "rds", "temp", "scup_survdat_25sims.rds"))
new_locs <- readRDS(here("data", "rds", "temp", "new_locations.rds"))


```

scup survey 
```{r}
plot_survey(scup_surv_sq, which_year = 1, which_sim = 1)
```


Extract fall survey locations of scup status quo survey 

```{r}
sq_sets <- scup_surv_sq$setdet |> 
  select(!c(N, n, n_measured, n_aged))

```

## SIMULATE SUMMER FLOUNDER SURVEY
### Status quo survey location
Fall distribution and survey 
```{r}
set.seed(131)
sf_survdat_sq <- sumflounder |> 
  sim_survey(n_sims = 25,
             trawl_dim = c(2.7, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5), #if plot = TRUE, xlim errors
             custom_sets = sq_sets, 
             set_den = 0.001,
             min_sets = 3,
             #lengths_cap = # max # of lengths to collect per set
             #age_cap = 8, # max # of ages to sample per length group - all 8?
             age_sampling = "stratified", 
             #age_length_group = , # length group bin size (cm) - 
             age_space_group = "set", # spatial scale of stratified age sampling - ages taken each tow?
             resample_cells = TRUE)
#saveRDS(sf_survdat_sq, here("R", "sumflounder", "data", "sumflounder_fall-sq-surv.rds"))

```

```{r}
plot_survey(sf_survdat_sq, which_year = 1, which_sim = 1)
```

### Preclusion 
Precluded summer flounder fall survey
```{r}
sf_survdat_precl <- sf_survdat_sq$setdet |> 
  filter(AREA_CODE == 2)

#saveRDS(sf_survdat_precl, here("R", "sumflounder", "data", "sumflounder_fall-precl-surv.rds"))
```

```{r}

sf_survdat_precl1 <-  sf_survdat_sq
sf_survdat_precl1[[15]] <- sf_survdat_precl1$setdet |> 
  filter(AREA_CODE == 2)
sf_survdat_precl1[[16]] <- sf_survdat_precl1$samp[sf_survdat_precl1$samp$set %in% sf_survdat_precl1$setdet$set, ]

 xax <- list(
    title = "",
    zeroline = FALSE,
    showline = FALSE,
    showticklabels = FALSE,
    showgrid = FALSE,
    ticks = ""
  )
  yax <- c(scaleanchor = "x", xax)

 #sp_strat <- raster::rasterToPolygons(sf_survdat_precl1$grid)#$strat)
  df_strat <- suppressMessages(ggplot2::fortify(strata)) #%>% group_by(.data$group))
  df_windareas <- suppressMessages(ggplot2::fortify(wind_areas_utm))

  setdet <- sf_survdat_precl1$setdet
  setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ]
  samp <- sf_survdat_precl1$samp
  samp <- samp[samp$set %in% setdet$set, ]
  samp <- merge(setdet[, c("year", "sim", "set", "x", "y")],
                samp, by = "set", all = TRUE)
  
 d <- crosstalk::SharedData$new(samp, ~set)

  base <- plot_ly(data = d)

  sp_p <- base %>%
    group_by(set) %>%
    summarise(x = unique(.data$x), y = unique(.data$y), n = sum(!is.na(.data$measured))) %>%
    add_markers(x = ~x, y = ~y, text = ~n,
                color = ~n, name = "n",
                showlegend = FALSE,
                marker = list(size = ~.scale_between(n, 2, 600),
                              sizemode = "area")) %>%
    add_paths(data = df_strat, x = ~long, y = ~lat, color = I("black"),
              hoverinfo = "none", size = I(0.5), showlegend = FALSE,
              alpha = 0.1) %>%
    add_paths(data = df_windareas, x = ~long, y = ~lat, color = I("black"),
              hoverinfo = "none", size = I(0.5), showlegend = FALSE,
              alpha = 0.1) %>%
    layout(xaxis = xax, yaxis = yax,
           margin = list(t = 0, r = 0, l = 0, b = 0, pad = 0))


```



Check how many tows were removed and that the amount is equal to the number of sets being reallocated. 
```{r}
sf_survdat_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year, sim, strat) |> 
  summarise(n_cells = length(cell), 
            n_sets = length(unique(set)))

scup_surv_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year, sim, strat) |> 
  summarise(n_cells = length(cell), 
            n_sets = length(unique(set)))

sf_survdat_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year) |> 
   summarise(n_sims = length(sim), 
             n_cells = length(cell), 
             n_sets = length(unique(set)))
scup_surv_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year) |> 
   summarise(n_sims = length(sim), 
             n_cells = length(cell), 
             n_sets = length(unique(set)))

scup_surv_sq$setdet |> filter(AREA_CODE == 2) |> nrow()
```


```{r}
ggplot() + 
  geom_sf(data = strata, fill = NA) +
  geom_sf(data = wind_areas_utm, fill = "lightblue", alpha = 0.5) + 
  geom_point(data = sf_survdat_precl, aes(x*1000, y *1000, color = n, size = n)) + 
  scale_color_continuous(type = "viridis") +
  scale_size_continuous(guide = "none") +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude", color = "Number") +
  theme(legend.position = "bottom", panel.spacing =  unit(1, "lines"), axis.text.x = element_text(angle = 45, vjust = 0.5))
```


### Reallocation
Reallocated survey effort 
```{r}
set.seed(123)
sf_survdat_newloc <- sumflounder |> 
  sim_survey(n_sims = 25,
             trawl_dim = c(2.7, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5), #if plot = TRUE, xlim errors
             custom_sets = new_locs, 
             set_den = 0.001,
             min_sets = 3,
             #lengths_cap = # max # of lengths to collect per set
             #age_cap = 8, # max # of ages to sample per length group - all 8?
             age_sampling = "stratified", 
             #age_length_group = , # length group bin size (cm) - 
             age_space_group = "set", # spatial scale of stratified age sampling - ages taken each tow?
             resample_cells = TRUE)
# saveRDS(sf_survdat_newloc, here("R", "sumflounder", "data", "sumflounder_fall-realloc-towdat.rds"))
```

```{r}
sf_survdat_realloc <- bind_rows(sf_survdat_precl, sf_survdat_newloc$setdet)
# saveRDS(sf_survdat_realloc, here("R", "sumflounder", "data", "sumflounder_fall-reallocated-surv.rds"))
```
** adding the reallocated survey locations to the precluded dataset does not generate the same number of observations as the status quo survey; reduced effort of 344 tows overall (how many each year, and simulation?)


```{r}
ggplot() + 
  geom_sf(data = strata, fill = NA) +
  geom_sf(data = wind_areas_utm, fill = "lightblue", alpha = 0.5) + 
  geom_point(data = sf_survdat_realloc, aes(x*1000, y *1000, color = n, size = n)) + 
  scale_color_continuous(type = "viridis") +
  scale_size_continuous(guide = "none") +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude", color = "Number") +
  theme(legend.position = "bottom", panel.spacing =  unit(1, "lines"), axis.text.x = element_text(angle = 45, vjust = 0.5))
```
 


