---
title: "Simulate multispecies"
author: "Angelia Miller"
date: "2023-07-25"
output: html_document
---

## Objective ####


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)
library(sf)

sseep.analysis.dat <- "C:/Users/amiller7/Documents/cinar-osse/sseep-analysis/data"

```

Data
```{r}
strata <- readRDS(here(sseep.analysis.dat, "rds", "active_strata_utm.rds"))
wind_areas <- readRDS(here(sseep.analysis.dat, "rds", "wind_areas_062022", "all_wind_areas_Jun2022.rds"))
wind_areas_utm <- st_transform(wind_areas, crs = 32618)

sumflounder <- readRDS(here("R", "sumflounder", "data", "sumflounder_fall-dist.rds"))
atlmackerel <- readRDS(here("R", "atlmackerel", "data", "atlmackerel_dist.rds"))

#scup survey data 
scup_surv_sq <- readRDS(here("data", "rds", "temp", "scup_survdat_25sims.rds"))
new_locs <- readRDS(here("data", "rds", "temp", "new_locations.rds"))


```

scup survey 
```{r}
plot_survey(scup_surv_sq, which_year = 1, which_sim = 1)
```


Extract fall survey locations of scup status quo survey 

```{r}
sq_sets <- scup_surv_sq$setdet |> 
  select(!c(N, n, n_measured, n_aged))

```

## SIMULATE SUMMER FLOUNDER SURVEY
### Status quo survey location
Fall distribution and survey 
```{r}
set.seed(131)
sf_survdat_sq <- sumflounder |> 
  sim_survey(n_sims = 25,
             trawl_dim = c(2.7, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5), #if plot = TRUE, xlim errors
             custom_sets = sq_sets, 
             set_den = 0.001,
             min_sets = 3,
             #lengths_cap = # max # of lengths to collect per set
             #age_cap = 8, # max # of ages to sample per length group - all 8?
             age_sampling = "stratified", 
             #age_length_group = , # length group bin size (cm) - 
             age_space_group = "set", # spatial scale of stratified age sampling - ages taken each tow?
             resample_cells = TRUE)
#saveRDS(sf_survdat_sq, here("R", "sumflounder", "data", "sumflounder_fall-sq-surv.rds"))

```

```{r}
plot_survey(sf_survdat_sq, which_year = 1, which_sim = 1)
```

### Preclusion 
Precluded summer flounder fall survey
```{r}
sf_survdat_precl <- sf_survdat_sq$setdet |> 
  filter(AREA_CODE == 2)

#saveRDS(sf_survdat_precl, here("R", "sumflounder", "data", "sumflounder_fall-precl-surv.rds"))
```


Check how many tows were removed and that the amount is equal to the number of sets being reallocated. 
```{r}
sf_survdat_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year, sim, strat) |> 
  summarise(n_cells = length(cell), 
            n_sets = length(unique(set)))

scup_surv_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year, sim, strat) |> 
  summarise(n_cells = length(cell), 
            n_sets = length(unique(set)))

sf_survdat_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year) |> 
   summarise(n_sims = length(sim), 
             n_cells = length(cell), 
             n_sets = length(unique(set)))
scup_surv_sq$setdet |> 
  filter(AREA_CODE == 1) |> 
  group_by(year) |> 
   summarise(n_sims = length(sim), 
             n_cells = length(cell), 
             n_sets = length(unique(set)))

scup_surv_sq$setdet |> filter(AREA_CODE == 2) |> nrow()
```


```{r}
ggplot() + 
  geom_sf(data = strata, fill = NA) +
  geom_sf(data = wind_areas_utm, fill = "lightblue", alpha = 0.5) + 
  geom_point(data = sf_survdat_precl, aes(x*1000, y *1000, color = n, size = n)) + 
  scale_color_continuous(type = "viridis") +
  scale_size_continuous(guide = "none") +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude", color = "Number") +
  theme(legend.position = "bottom", panel.spacing =  unit(1, "lines"), axis.text.x = element_text(angle = 45, vjust = 0.5))
```


### Reallocation
Reallocated survey effort 
```{r}
set.seed(123)
sf_survdat_newloc <- sumflounder |> 
  sim_survey(n_sims = 25,
             trawl_dim = c(2.7, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5), #if plot = TRUE, xlim errors
             custom_sets = new_locs, 
             set_den = 0.001,
             min_sets = 3,
             #lengths_cap = # max # of lengths to collect per set
             #age_cap = 8, # max # of ages to sample per length group - all 8?
             age_sampling = "stratified", 
             #age_length_group = , # length group bin size (cm) - 
             age_space_group = "set", # spatial scale of stratified age sampling - ages taken each tow?
             resample_cells = TRUE)
# saveRDS(sf_survdat_newloc, here("R", "sumflounder", "data", "sumflounder_fall-realloc-towdat.rds"))
```

```{r}
sf_survdat_realloc <- bind_rows(sf_survdat_precl, sf_survdat_newloc$setdet)
# saveRDS(sf_survdat_realloc, here("R", "sumflounder", "data", "sumflounder_fall-reallocated-surv.rds"))
```
** adding the reallocated survey locations to the precluded dataset does not generate the same number of observations as the status quo survey; reduced effort of 344 tows overall (how many each year, and simulation?)


```{r}
ggplot() + 
  geom_sf(data = strata, fill = NA) +
  geom_sf(data = wind_areas_utm, fill = "lightblue", alpha = 0.5) + 
  geom_point(data = sf_survdat_realloc, aes(x*1000, y *1000, color = n, size = n)) + 
  scale_color_continuous(type = "viridis") +
  scale_size_continuous(guide = "none") +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude", color = "Number") +
  theme(legend.position = "bottom", panel.spacing =  unit(1, "lines"), axis.text.x = element_text(angle = 45, vjust = 0.5))
```
 


