---
title: "Sim Survey Trials Sandbox"
author: "Angelia Miller & Catalina Roman"
date: "2023-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(here)
```

# Summer flounder stock assessment data 
```{r}
## 66th SAW Report
Rec_age0 <- c(48689, 60598, 44582, 33088, 28416) #most recent 5 years
mean(Rec_age0)
sd(Rec_age0)

F <- c(0.340, 0.286, 0.331, 0.414, 0.419) #most recent 5 years
M <- 0.2

Z <- F + M

mean(Z)
sd(Z)
log(sd(Z))
```


# Dummy Grid
With Summer Flounder stock assessment data and NEFSC Trawl dimensions and selectivity
```{r}

#with temporary grid
g <- make_grid(res = c(10, 10), n_div = 4, strat_splits = 2,
               shelf_depth = 100, shelf_width = 80, depth_range = c(10, 375))

pop_dist1 <- sim_abundance(ages = 0:7, years = 1:10,
                           R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                           Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                           growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) |>
             sim_distribution(grid = g,
                              ays_covar = sim_ays_covar(range = 89.5),
                              depth_par = sim_parabola(mu = 71.79,
                                                       sigma = 52.92))

# pop_dist$N<- pop_dist$N* (10^5)
# pop_dist$N_at_length <- pop_dist$N_at_length * (10^8)


pop_dist1 |>
  sim_survey(n_sims = 2,
             trawl_dim = c(0.014, 0.19),
             q = sim_logistic(k = 2, x0 = 2.5))

```


# New foundland Grid
With Summer Flounder stock assessment data and NEFSC Trawl dimensions and selectivity
```{r}
survey_grid <- survey_grid #save package grid into R environment

pop_dist2 <- sim_abundance(ages = 0:7, years = 1:10,
                           R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                           Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                           growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) |>
             sim_distribution(grid = survey_grid,
                              ays_covar = sim_ays_covar(range = 89.5),
                              depth_par = sim_parabola(mu = 71.79,
                                                       sigma = 52.92))

pop_dist2 |>
  sim_survey(n_sims = 2,
             trawl_dim = c(0.014, 0.19),
             q = sim_logistic(k = 2, x0 = 2.5))

```

# With sdmTMB grid
```{r}
# load the sdmtmb grid
sdmtmb_grid <- readRDS(file = here(sdmtmb.dir, "data", "survey_grid.rds")) |>
  rename(strat = STRATUM) |>
  mutate(division = 1, 
         cell = seq(1:length(depth)), 
         depth = abs(depth), 
         X = X/1000, 
         Y = Y/1000) |>
  dplyr::select(X, Y, cell, division, strat, depth)

# CONVERT GRID 
# to raster
sdmtmb_ras <- rasterFromXYZ(sdmtmb_grid, crs = 32618)
#then to RasterStack
sdmtmb_stck <- stack(sdmtmb_ras)
#writeRaster(sdmtmb_ras, filename = here("data", "survey_grid.grd"))

pop_dist3 <- sim_abundance(ages = 0:7, years = 1:10,
                           R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                           Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                           growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) |>
             sim_distribution(grid = sdmtmb_stck,
                              ays_covar = sim_ays_covar(range = 89.5),
                              depth_par = sim_parabola(mu = 71.79,
                                                       sigma = 52.92))

pop_dist3 |>
  sim_survey(n_sims = 2,
             trawl_dim = c(0.014, 0.19),
             q = sim_logistic(k = 2, x0 = 2.5))
```



# Hardwire Summer flounder sdmTMB data 
## Set up the Data
```{r}
#### LOAD DATA ####
# load summer flounder data
#sumflounder <- readRDS(file = here(sdmtmb.dir, "data", "sumflounder.rds"))

# load spring model
#m7_spring <- readRDS(file = here(sdmtmb.dir, "model-outputs", "m7_spring.rds"))

# load spring model predictions
spring_preds <- readRDS(file = here(sdmtmb.dir, "data", "spring_preds.rds"))

# filter distribution predictions for last 5 years
preds_5yr<- spring_preds |>
  filter(EST_YEAR %in% c(2017:2021)) |>
  rename(N_dist = est,
         year = EST_YEAR) |>
  dplyr::select(c(X,Y, year, N_dist)) |>
  mutate(year = case_when(
    year == 2017 ~ 1, 
    year == 2018 ~ 2, 
    year == 2019 ~ 3, 
    year == 2020 ~ 4, 
    year == 2021 ~ 5
  ),
         cell = seq(1:length(N_dist))) |>
  data.table::as.data.table()
```

## Simulate Abundance 
```{r}
set.seed(8675309)
pop <- sim_abundance(ages = 0:7, years = 1:5,
                     R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                     Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                     growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) #66th SAW Report ASFMC
Nage <- tibble(age = 0:7, pop$N0)
plot_surface(pop, mat = "N")

```

## Append the simulated abundance 
```{r}
# append  to abundance and rename list item
pop <- append(pop,sdmtmb_stck) # use RasterStack object of sdmtmb grid
names(pop)[[10]] <- "grid"

# resave dataframe grid to match dummy data grid
grid_xy <- sdmtmb_grid |>
 # dplyr::select(X, Y, depth, division, strat) |>
  rename(x = X, y = Y)# |>
  #mutate(cell = seq(1:length(depth)))

# append to the simulated abundance object
pop <- append(pop, list(grid_xy))
names(pop)[[11]] <- "grid_xy"

# calculate numbers at age in each cell based on the year distribution
dist <- sdmTMB::replicate_df(preds_5yr, "age", c(0:7)) |>
  left_join(Nage, by = "age") |>
  rename(Nage = "pop$N0") |>
  mutate(N_dist = exp(N_dist),
         P_i = N_dist/sum(N_dist),
         N = Nage * P_i) |>
  dplyr::select(age, year, cell, N, X, Y)

# append to the simulated abundance object
pop <- append(pop, list(dist))
names(pop)[[12]] <- "sp_N"

# plot numbers at age
ggplot(pop$sp_N) + geom_tile(aes(X, Y, fill = N), width = 10, height = 10) + scale_fill_viridis_c() + facet_wrap(~age)
```

## Simulate Survey
```{r}

#a 4-seam, 3-bridle net using a rock-hopper sweep aboard the FSV Henry B. Bigelow since 2009
#lo_effort <- 
pop |>
  sim_survey(n_sims = 2,
             trawl_dim = c(0.014, 0.19),
             q = sim_logistic(k = 2, x0 = 2.5))


# pop$sp_N$N <- round(pop$sp_N$N)
# N <- tapply(pop$sp_N$N, list(pop$sp_N$age, pop$sp_N$year), sum)
# N <- N[rownames(pop$N), colnames(pop$N)]
# names(N) <- names(pop$N)
# pop$N_at_length <- convert_N(N_at_age = N,
#                              lak = pop$sim_length(age = pop$ages, length_age_key = TRUE))
# pop$N <- N
# pop$N0 <- N[[1]] #want age 0
# pop$R <- N[[1]] # want value at age 0
# pop
             #set_den = 0.0005)
# hi_effort <- pop_dist |>
#   sim_survey(set_den = 0.0025)
#
# plot_survey(lo_effort, which_year = 1, which_sim = 1)
# plot_survey(hi_effort, which_year = 1, which_sim = 1)
```





