---
title: "Sim Survey Trials Sandbox"
author: "Angelia Miller & Catalina Roman"
date: "2023-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)

sdmtmb.dir <- "../sseep-analysis/sdmtmb"
sseep.analysis <- "../sseep-analysis"
```

# Summer flounder stock assessment data 
```{r}
set.seed(8675309)

## 66th SAW Report
Rec_age0 <- c(48689, 60598, 44582, 33088, 28416) #most recent 5 years
mean(Rec_age0)
sd(Rec_age0)

F <- c(0.340, 0.286, 0.331, 0.414, 0.419) #most recent 5 years
M <- 0.2

Z <- F + M

mean(Z)
sd(Z)
log(sd(Z))
```


# Dummy Grid
With Summer Flounder stock assessment data and NEFSC Trawl dimensions and selectivity
```{r}

#with temporary grid
g <- make_grid(res = c(10, 10), n_div = 4, strat_splits = 2,
               shelf_depth = 100, shelf_width = 80, depth_range = c(10, 375))

pop_dist1 <- sim_abundance(ages = 0:7, years = 1:10,
                           R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                           Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                           growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) |>
             sim_distribution(grid = g,
                              ays_covar = sim_ays_covar(range = 89.5),   # value pulled from Summer flounder sdmtmb model
                              depth_par = sim_parabola(mu = 71.79,       # value pulled from Summer flounder sdmtmb model
                                                       sigma = 52.92))   # value pulled from Summer flounder sdmtmb model

# pop_dist$N<- pop_dist$N* (10^5)
# pop_dist$N_at_length <- pop_dist$N_at_length * (10^8)


pop_dist1 |>
  sim_survey(n_sims = 2,
             trawl_dim = c(0.014, 0.19),
             q = sim_logistic(k = 2, x0 = 2.5))

```


# New foundland Grid
With Summer Flounder stock assessment data and NEFSC Trawl dimensions and selectivity
```{r}
survey_grid <- SimSurvey::survey_grid #save package grid into R environment

pop_dist2 <- sim_abundance(ages = 0:7, years = 1:10,
                           R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                           Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                           growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) |>
             sim_distribution(grid = survey_grid,
                              ays_covar = sim_ays_covar(range = 89.5),  # value pulled from Summer flounder sdmtmb model
                              depth_par = sim_parabola(mu = 71.79,      # value pulled from Summer flounder sdmtmb model
                                                       sigma = 52.92))  # value pulled from Summer flounder sdmtmb model

profvis::profvis(pop_dist2 |>
  sim_survey(n_sims = 2,
             trawl_dim = c(0.012, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5)))

```

# With sdmTMB grid
```{r}
# load the sdmtmb grid
grid <- readRDS(file = here(sdmtmb.dir, "data", "survey_grid.rds")) |>
  rename(strat = STRATUM, 
         depth = AVGDEPTH) |>
  mutate(division = 1, 
         x = X/1000, 
         y = Y/1000) |>
  dplyr::select(x, y, cell, division, strat, depth) |>
  as.data.frame()

crs <-  "+proj=utm +zone=18 +datum=WGS84 +units=km +no_defs"

grid_ras <- rasterFromXYZ(grid, crs = crs)
grid_stck <- stack(grid_ras)

pop_dist3 <- sim_abundance(ages = 0:7, years = 1:10,
                           R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                           Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                           growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) |>
             sim_distribution(grid = grid_stck,
                              ays_covar = sim_ays_covar(range = 89.5),  # value pulled from Summer flounder sdmtmb model
                              depth_par = sim_parabola(mu = 71.79,      # value pulled from Summer flounder sdmtmb model
                                                       sigma = 52.92))  # value pulled from Summer flounder sdmtmb model

surv_dat <- pop_dist3 |>
  sim_survey(n_sims = 2,
             trawl_dim = c(1.85, 0.014), # distance towed, trawl width; km
             q = sim_logistic(k = 2, x0 = 2.5), 
             resample_cells = TRUE, 
             set_den = 0.0025)
# htmlwidgets::saveWidget(p, "profile.html")
# browseURL("profile.html")
```

## Calculate Stratified Mean Abundance Index 
```{r}
# Load Extra Data 
strata_wts <- readRDS(here(sseep.analysis, "data", "rds", "strata_wts.rds")) |> 
  rename(strat = STRATUM)

data <- surv_dat$setdet |> as_tibble()

# check individual strata information
x <- data |>
    filter(strat==1010, year==1, sim==1)

individual <- data |>
    group_by(strat, year, sim) |>
    dplyr::summarise(towct = length(set), # calculate unique tows
              mu = sum(N)/towct, # find the average biomass based on unique tows rather than observations to avoid potential duplication
              var = ifelse(towct == 1, 0, # if the tow count equals 1, then variance about the mean should be 0
                           sum((N - mu)^2)/(towct - 1)), .groups = "keep") # if tow count does not equal 1, then find the variance of biomass
weighted <- individual |>
    left_join(strata_wts, by = "strat") |> # add each stratum area and relative weight to the dataset based on STRATUM number
      group_by(strat, year, sim) |>
    mutate(wt_mu = Area_SqNm * mu, # part one of the stratified mean formula
           wt_var = ((((RelWt)^2) * var) / towct) * (1 - (towct / Area_SqNm)), 
           wt_var = ifelse(is.na(wt_var), 0, wt_var)) 
           # part one of the stratified variance formula

BTSArea <- as.integer(sum(strata_wts$Area_SqNm))

stratified <- weighted |>
    group_by(year, sim) |>
    summarise(stratmu = (sum(wt_mu)) / BTSArea, # part two of the stratified mean formula
              stratvar = sum(wt_var), .groups = "keep") |>
  mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), #logistic standard deviation
         lower = qlnorm(0.025, log(stratmu), sdlog), # lower quantile of the logistic normal distribution
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> # upper quantile of the logistic normal distribution
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), # if sdlog is NaN, replace with 0
         lower = ifelse(is.nan(lower), 0, lower), # if the lower quantile is NaN, replace with 0
         upper = ifelse(is.nan(upper), 0, upper))


sf_stratmu_obs <- readRDS(here("data", "rds", "num-strat-mu_all.rds")) |> 
  filter(SVSPP == 103, TYPE == "With Wind Included") |> 
   mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), #logistic standard deviation
         lower = qlnorm(0.025, log(stratmu), sdlog), # lower quantile of the logistic normal distribution
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> # upper quantile of the logistic normal distribution
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), # if sdlog is NaN, replace with 0
         lower = ifelse(is.nan(lower), 0, lower), # if the lower quantile is NaN, replace with 0
         upper = ifelse(is.nan(upper), 0, upper))



sim_plot <- ggplot() +
  #geom_point() +
  geom_pointrange(data = stratified, aes(x = as.factor(year), y = stratmu, color = as.factor(sim), ymin=lower, ymax = upper), position = position_dodge2(width=0.4)) +
  #facet_wrap(vars(SEASON), scales = "free_y") +
  #facet_grid(rows = vars(GEO_AREA), cols = vars(SEASON), scales = "free_y") + 
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "Simulation") +
  #ylim(0,NA) +
  theme_bw() #+ 

obs_plot <-  ggplot() +
  geom_pointrange(data = sf_stratmu_obs, aes(x = as.factor(EST_YEAR), y = stratmu, color = as.factor(SEASON), ymin=lower, ymax = upper), position =  position_dodge2(width=0.4)) +
  #facet_wrap(vars(SEASON), scales = "free_y") +
  #facet_grid(rows = vars(GEO_AREA), cols = vars(SEASON), scales = "free_y") + 
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "SEASON") +
  #ylim(0,NA) +
  theme_bw()

sim_plot + obs_plot


```


# Hardwire Summer flounder sdmTMB data 
## Set up the Data
```{r}
#### LOAD DATA ####
#sdmtmb.dir <- "../sseep-analysis/sdmtmb"

# load spring model predictions
spring_preds <- readRDS(file = here(sdmtmb.dir, "data", "spring_projects.rds"))

# filter distribution predictions for last 5 years
preds_5yr<- spring_preds |>
  filter(EST_YEAR %in% c(2017:2021)) |>
  rename(year = EST_YEAR) |>
   mutate(N_dist = exp(est), 
         year = case_when(
    year == 2017 ~ 1, 
    year == 2018 ~ 2, 
    year == 2019 ~ 3, 
    year == 2020 ~ 4, 
    year == 2021 ~ 5
  ),
         cell = seq(1:length(N_dist))) |>
  dplyr::select(c(X,Y, year, N_dist, cell)) |>
  data.table::as.data.table()
```

## Simulate Abundance 
```{r}
pop <- sim_abundance(ages = 0:7, years = 1:5,
                     R = sim_R(log_mean = log(mean(Rec_age0)), log_sd = 0.001, plot = TRUE),
                     Z = sim_Z(log_mean = log(mean(Z)), log_sd = 0.001, plot = TRUE),
                     growth = sim_vonB(Linf = 83.6, K = 0.14, plot = TRUE)) #66th SAW Report ASFMC
Nage <- tibble(age = 0:7, pop$N0)
plot_surface(pop, mat = "N")

```

## Append the simulated abundance 
```{r}
# append  to abundance and rename list item
pop <- append(pop,grid_stck) # use RasterStack object of sdmtmb grid
names(pop)[[10]] <- "grid"

# resave dataframe grid to match dummy data grid
grid_xy <- grid |>
 # dplyr::select(X, Y, depth, division, strat) |>
  #rename(x = X, y = Y) |>
  data.table::as.data.table()
  #mutate(cell = seq(1:length(depth)))

# append to the simulated abundance object
pop <- append(pop, list(grid_xy))

names(pop)[[11]] <- "grid_xy"

# calculate numbers at age in each cell based on the year distribution
dist <- sdmTMB::replicate_df(preds_5yr, "age", c(0:7)) |>
  left_join(Nage, by = "age") |>
  rename(Nage = "pop$N0", 
         x = X, 
         y = Y) |>
  mutate(P_i = N_dist/sum(N_dist),
         N = Nage * P_i, 
         age = as.double(age), 
         cell = as.double(cell)) |>
  dplyr::select(age, year, cell, N, x, y)

# append to the simulated abundance object
pop <- append(pop, list(dist))
names(pop)[[12]] <- "sp_N"

# plot numbers at age
ggplot(pop$sp_N) + geom_tile(aes(x, y, fill = N), width = 10, height = 10) + scale_fill_viridis_c() + facet_wrap(~age)
```

## Simulate Survey
```{r}

#a 4-seam, 3-bridle net using a rock-hopper sweep aboard the FSV Henry B. Bigelow since 2009
#lo_effort <- 
surv_dat2 <- pop |>
  sim_survey(n_sims = 2,
             trawl_dim = c(1.85, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5),
             resample_cells = TRUE)

surv_dat3 <- pop |>
  sim_survey(n_sims = 2,
             trawl_dim = c(1.85, 0.014),
             q = sim_logistic(k = 2, x0 = 2.5), 
             set_den = 0.02,
             resample_cells = TRUE)


# pop$sp_N$N <- round(pop$sp_N$N)
# N <- tapply(pop$sp_N$N, list(pop$sp_N$age, pop$sp_N$year), sum)
# N <- N[rownames(pop$N), colnames(pop$N)]
# names(N) <- names(pop$N)
# pop$N_at_length <- convert_N(N_at_age = N,
#                              lak = pop$sim_length(age = pop$ages, length_age_key = TRUE))
# pop$N <- N
# pop$N0 <- N[[1]] #want age 0
# pop$R <- N[[1]] # want value at age 0
# pop
             #set_den = 0.0005)
# hi_effort <- pop_dist |>
#   sim_survey(set_den = 0.0025)
#
# plot_survey(lo_effort, which_year = 1, which_sim = 1)
# plot_survey(hi_effort, which_year = 1, which_sim = 1)
```




```{r}
data <- surv_dat2$setdet |> as_tibble()

# check individual strata information
x <- data |>
    filter(strat==1010, year==1, sim==1)

individual <- data |>
    group_by(strat, year, sim) |>
    dplyr::summarise(towct = length(set), # calculate unique tows
              mu = sum(N)/towct, # find the average biomass based on unique tows rather than observations to avoid potential duplication
              var = ifelse(towct == 1, 0, # if the tow count equals 1, then variance about the mean should be 0
                           sum((N - mu)^2)/(towct - 1)), .groups = "keep") # if tow count does not equal 1, then find the variance of biomass
weighted <- individual |>
    left_join(strata_wts, by = "strat") |> # add each stratum area and relative weight to the dataset based on STRATUM number
      group_by(strat, year, sim) |>
    mutate(wt_mu = Area_SqNm * mu, # part one of the stratified mean formula
           wt_var = ((((RelWt)^2) * var) / towct) * (1 - (towct / Area_SqNm)), 
           wt_var = ifelse(is.na(wt_var), 0, wt_var)) 
           # part one of the stratified variance formula

BTSArea <- as.integer(sum(strata_wts$Area_SqNm))

stratified <- weighted |>
    group_by(year, sim) |>
    summarise(stratmu = (sum(wt_mu)) / BTSArea, # part two of the stratified mean formula
              stratvar = sum(wt_var), .groups = "keep") |>
  mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), #logistic standard deviation
         lower = qlnorm(0.025, log(stratmu), sdlog), # lower quantile of the logistic normal distribution
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> # upper quantile of the logistic normal distribution
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), # if sdlog is NaN, replace with 0
         lower = ifelse(is.nan(lower), 0, lower), # if the lower quantile is NaN, replace with 0
         upper = ifelse(is.nan(upper), 0, upper))


sf_stratmu_obs <- readRDS(here("data", "rds", "num-strat-mu_all.rds")) |> 
  filter(SVSPP == 103, TYPE == "With Wind Included") |> 
   mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), #logistic standard deviation
         lower = qlnorm(0.025, log(stratmu), sdlog), # lower quantile of the logistic normal distribution
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> # upper quantile of the logistic normal distribution
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), # if sdlog is NaN, replace with 0
         lower = ifelse(is.nan(lower), 0, lower), # if the lower quantile is NaN, replace with 0
         upper = ifelse(is.nan(upper), 0, upper))



sim_plot <- ggplot() +
  #geom_point() +
  geom_pointrange(data = stratified, aes(x = as.factor(year), y = stratmu, color = as.factor(sim), ymin=lower, ymax = upper), position = position_dodge2(width=0.4)) +
  #facet_wrap(vars(SEASON), scales = "free_y") +
  #facet_grid(rows = vars(GEO_AREA), cols = vars(SEASON), scales = "free_y") + 
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "Simulation") +
  #ylim(0,NA) +
  theme_bw() #+ 

obs_plot <-  ggplot() +
  geom_pointrange(data = sf_stratmu_obs |> filter(EST_YEAR %in% c(2017:2021)), aes(x = as.factor(EST_YEAR), y = stratmu, color = as.factor(SEASON), ymin=lower, ymax = upper), position =  position_dodge2(width=0.4)) +
  #facet_wrap(vars(SEASON), scales = "free_y") +
  #facet_grid(rows = vars(GEO_AREA), cols = vars(SEASON), scales = "free_y") + 
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "SEASON") +
  #ylim(0,NA) +
  theme_bw()

sim_plot + obs_plot
```


```{r}
# sdmtmb_grid <- readRDS(file = here(sdmtmb.dir, "data", "survey_grid.rds")) |>
#   rename(strat = STRATUM) |>
#   mutate(division = 1)
```



