---
title: "Strat mean"
date: "2024-01-07"
output: html_document
---

## Objective
true and estimated stratified abundance indices, bias and precision of those abundance indices


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(SimSurvey)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(patchwork)
library(here)

#sseep.analysis <- "C:/Users/croman1/Desktop/UMassD/sseep-analysis/"
```


### Load Data 
```{r}
# read in survey data 
# Status quo
surv_dat1 <- readRDS(here("R", "scup", "data", "scup_survdat_25sims_spr.rds"))
# Preclusion 
survey_preclusion <- surv_dat1
survey_preclusion$setdet <- filter(surv_dat1$setdet, AREA_CODE != 1) #filter wind areas

# Reallocation
surv_reallocation <- readRDS(here("R", "scup", "data", "scup_survreallocation_25sims_spr.rds"))
setdet_sr <- surv_reallocation$setdet 
setdet_sp <- survey_preclusion$setdet |>
  relocate(division, .after ="cell")


survey_setdet_reallocation <- setdet_sp |>
  bind_rows(setdet_sr)



# read in the area weights for each strata 
strata_wts <- readRDS(here("R","scup", "data","active_strata_wts.rds")) |>
  rename(strat = STRATUM)

# find the total survey area 
BTSArea <- as.integer(sum(strata_wts$Area_SqNm))

# extract survey grid datatable
grid_xy <- surv_dat1$grid_xy

obs_stratmu <- readRDS(here("R","scup","data", "scup_nowind_stratmu.rds")) |> 
  filter(EST_YEAR %in% c(2018,2019,2020,2021), SEASON == "SPRING")

```

Load data for maps
```{r, include=FALSE}
# read the strata shapefile in to plot the polygons based on proportion and overlay the tow points
strata <- readRDS(here("R", "scup","data", "active_strata.rds")) 

# wind areas 
wind_areas <- readRDS(here("R", "scup","data", "all_wind_areas_Jun2022.rds")) 
wind_areas2 <- readRDS(here("R", "scup", "data","all_wind_areas_Jan2023.rds")) 

coastline <- sf::st_read(dsn = here("R", "scup","data", "gis", "eastern_coast_UTM.shp"))

```


## Stratified mean from status quo survey
Bias - variance
precision - CV = sd/mu
```{r}
# pull out the simulated numbers of scup from the survey
est_N_s1 <- surv_dat1$setdet |> as_tibble()


# calculate the individual mean abundance in each strata for each year and simulation
estN_ind_s1 <- est_N_s1 |>
    group_by(strat, year, sim) |>
    dplyr::summarise(towct = length(set), 
              mu = sum(n)/towct, 
              var = ifelse(towct == 1, 0, 
                           sum((n - mu)^2)/(towct - 1)), .groups = "keep") 

# calculate the weighted average by multiplying by each stratas proportional area
estN_wtd_s1 <- estN_ind_s1 |>
    left_join(strata_wts, by = "strat") |> # add each stratum area and relative weight to the dataset based on STRATUM number
      group_by(strat, year, sim) |>
    mutate(wt_mu = Area_SqNm * mu, # part one of the stratified mean formula
           wt_var = ((((RelWt)^2) * var) / towct) * (1 - (towct / Area_SqNm)), 
           wt_var = ifelse(is.na(wt_var), 0, wt_var)) 
           # part one of the stratified variance formula

# calculated the stratified average and the standard deviations 
estN_strat_s1 <- estN_wtd_s1 |>
    group_by(year, sim) |>
    summarise(stratmu = (sum(wt_mu)) / BTSArea, 
              stratvar = sum(wt_var), .groups = "keep") |>
  mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), 
         lower = qlnorm(0.025, log(stratmu), sdlog), 
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> 
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), 
         lower = ifelse(is.nan(lower), 0, lower), 
         upper = ifelse(is.nan(upper), 0, upper), 
         TYPE = "Status quo")

```


Plot of strat mean over time for each simulation
```{r}
# create a plot of the indices of abundance from the simulated data 
estN_plot_s1 <- ggplot() +
  geom_pointrange(data = estN_strat_s1, aes(x = as.factor(year), 
                                         y = stratmu, color = as.factor(sim), 
                                         ymin=lower, ymax = upper), 
                  position = position_dodge2(width=0.4)) +
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "Simulation", 
       title = "Stratified mean",
       subtitle = "Status Quo Survey") +
  #ylim(0,30)+
  theme_bw() 

estN_plot_s1
```

#Map of survey placements
```{r}
#status quo survey
est_N_s1
#convert to dec deg
est_N_s1_dec <- utm2lonlat(est_N_s1$x,est_N_s1$y, zone=18, hemisphere = "N", km=TRUE)
est_N_s1 <- est_N_s1 |> 
  bind_cols(est_N_s1_dec)


(sq_map <-  ggplot(strata) + 
  geom_sf() +  
 # geom_sf(data = coastline, fill = "grey70", color = "black") + 
  geom_sf(data = wind_areas, fill="darkblue", color="grey80") +
  geom_point(data=est_N_s1, aes(x = longitude, y = latitude), 
             color = "darkorange2", shape = 19, alpha=0.2, size=0.001)  +
  labs(title = "", x = "Long", y = "Lat"))

```


##Stratified mean from survey with precluded tows
```{r}
# pull out the simulated numbers of scup from the survey with precluded data inside wind areas
est_N_s2 <- survey_preclusion$setdet |> as_tibble()


# calculate the individual mean abundance in each strata for each year and simulation
estN_ind_s2 <- est_N_s2 |>
    group_by(strat, year, sim) |>
    dplyr::summarise(towct = length(set), 
              mu = sum(n)/towct, 
              var = ifelse(towct == 1, 0, 
                           sum((n - mu)^2)/(towct - 1)), .groups = "keep") 

# calculate the weighted average by multiplying by each stratas proportional area
estN_wtd_s2 <- estN_ind_s2 |>
    left_join(strata_wts, by = "strat") |>
      group_by(strat, year, sim) |>
    mutate(wt_mu = Area_SqNm * mu, 
           wt_var = ((((RelWt)^2) * var) / towct) * (1 - (towct / Area_SqNm)), 
           wt_var = ifelse(is.na(wt_var), 0, wt_var)) 
           

# calculated the stratified average and the standard deviations 
estN_strat_s2 <- estN_wtd_s2 |>
    group_by(year, sim) |>
    summarise(stratmu = (sum(wt_mu)) / BTSArea, 
              stratvar = sum(wt_var), .groups = "keep") |>
  mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), 
         lower = qlnorm(0.025, log(stratmu), sdlog), 
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> 
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), 
         lower = ifelse(is.nan(lower), 0, lower), 
         upper = ifelse(is.nan(upper), 0, upper), 
         TYPE = "Preclusion")

```


Plot of strat mean over time for each simulation
```{r}

estN_plot_s2 <- ggplot() +
  geom_pointrange(data = estN_strat_s2, aes(x = as.factor(year), 
                                         y = stratmu, color = as.factor(sim), 
                                         ymin=lower, ymax = upper), 
                  position = position_dodge2(width=0.4)) +
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "Simulation", 
       title = "Stratified mean",
       subtitle = "Preclusion") +
  #ylim(0,25) +
  theme_bw() 

estN_plot_s2
```

#Map of survey placements
```{r}
#preclusion scenario survey
est_N_s2
#convert to dec deg
est_N_s2_dec <- utm2lonlat(est_N_s2$x,est_N_s2$y, zone=18, hemisphere = "N", km=TRUE)
est_N_s2 <- est_N_s2 |> 
  bind_cols(est_N_s2_dec)

(s2_map <-  ggplot(strata) + 
  geom_sf() +  
 # geom_sf(data = coastline, fill = "grey70", color = "black") + 
  geom_sf(data = wind_areas, fill="darkblue", color="grey80") +
  geom_point(data=est_N_s2, aes(x = longitude, y = latitude), 
             color = "darkorange3", shape = 20, alpha=0.2, size=0.01)  +
  labs(title = "", x = "Long", y = "Lat"))


```


## ##Stratified mean from survey with reallocated tows
```{r}
est_N_s3 <- survey_setdet_reallocation |> as_tibble()


# calculate the individual mean abundance in each strata for each year and simulation
estN_ind_s3 <- est_N_s3 |>
    group_by(strat, year, sim) |>
    dplyr::summarise(towct = length(set), 
              mu = sum(n)/towct, 
              var = ifelse(towct == 1, 0, 
                           sum((n - mu)^2)/(towct - 1)), .groups = "keep") 

# calculate the weighted average by multiplying by each stratas proportional area
estN_wtd_s3 <- estN_ind_s3 |>
    left_join(strata_wts, by = "strat") |>
      group_by(strat, year, sim) |>
    mutate(wt_mu = Area_SqNm * mu, 
           wt_var = ((((RelWt)^2) * var) / towct) * (1 - (towct / Area_SqNm)), 
           wt_var = ifelse(is.na(wt_var), 0, wt_var)) 
           

# calculated the stratified average and the standard deviations 
estN_strat_s3 <- estN_wtd_s3 |>
    group_by(year, sim) |>
    summarise(stratmu = (sum(wt_mu)) / BTSArea, 
              stratvar = sum(wt_var), .groups = "keep") |>
  mutate(sdlog = sqrt(log(1+(sqrt(stratvar)/stratmu)^2)), 
         lower = qlnorm(0.025, log(stratmu), sdlog), 
         upper = qlnorm(0.975, log(stratmu), sdlog)) |> 
  mutate(sdlog = ifelse(is.nan(sdlog), 0, sdlog), 
         lower = ifelse(is.nan(lower), 0, lower), 
         upper = ifelse(is.nan(upper), 0, upper), 
         TYPE = "Reallocation")

```


Plot of strat mean over time for each simulation
```{r}

estN_plot_s3 <- ggplot() +
  geom_pointrange(data = estN_strat_s3, aes(x = as.factor(year), 
                                         y = stratmu, color = as.factor(sim), 
                                         ymin=lower, ymax = upper), 
                  position = position_dodge2(width=0.4)) +
  labs(x = "YEAR", y = "Stratified Mean (N/tow)", color = "Simulation", 
       title = "Stratified mean",
       subtitle = "Reallocation") +
 # ylim(0,25)+
  theme_bw() 

estN_plot_s3
```

#Map of survey placements
```{r}
#reallocation survey
setdet_sr
setdet_sp
#convert to dec deg
setdet_sr_dec <- utm2lonlat(setdet_sr$x,setdet_sr$y, zone=18, hemisphere = "N", km=TRUE)
setdet_sr <- setdet_sr |> 
  bind_cols(setdet_sr_dec)

setdet_sp_dec <- utm2lonlat(setdet_sp$x,setdet_sp$y, zone=18, hemisphere = "N", km=TRUE)
setdet_sp <- setdet_sp |> 
  bind_cols(setdet_sp_dec)

(s3_map <-  ggplot(strata) + 
  geom_sf() +  
 # geom_sf(data = coastline, fill = "grey70", color = "black") + 
  geom_sf(data = wind_areas, fill="lightcyan1", color="grey80") +
  geom_point(data=setdet_sp, aes(x = longitude, y = latitude), 
             color = "darkorange3", shape = 20, alpha=0.2, size=0.01)  +
      geom_point(data=setdet_sr, aes(x = longitude, y = latitude), 
             color = "royalblue", shape = 20, alpha=0.2, size=1)  +
  labs(title = "", x = "Long", y = "Lat"))
```





pendant
Mean Standard Error
```{r}
estN_ind_s1

```

## Error 
```{r}
# average sampling error for mean catch rate 
estN_se_s1 <- estN_ind_s1 |>
  mutate(sd = sqrt(var), 
         se = (sd / sqrt(towct))*1.96)

avg_se_s1 <- estN_se_s1 |> 
  group_by(strat, year, sim) |> 
  summarise(avg_se = mean(se))

avg_se_s1


# error statistics for abundance index using the error_stats() in SimSurvey
observed <- obs_stratmu |>
  ungroup() |>
  mutate(year = case_when( # change the year values based on their sequence 
         EST_YEAR == 2016 ~ 1,
         EST_YEAR == 2018 ~ 2,
         EST_YEAR == 2019 ~ 3,
         EST_YEAR == 2021 ~ 4
         )) |>
  select(year, stratmu, stratvar) |>
  rename(I = stratmu, 
         obs_stratvar = stratvar)

comp_1 <- estN_strat_s1 |> 
  select(year, sim, stratmu, stratvar) |> 
  rename(I_hat = stratmu,
         est_stratvar = stratvar) |> 
  left_join(observed, by = "year")

comp_1$error <- comp_1$I_hat - comp_1$I # also bias

means_1 <- error_stats(comp_1$error)

means_1

```


## Precision
 - "the precision of the mean density estimate of species g
at time t across replicate surveys and is the standard deviation of the simulated survey estimates (where rgt is the mean density estimate of species g at time t averaged across the D surveys) relative to ygt , the true mean density of species g at time t" (Oyafuso et al 2021)
 - "The True CV (TrueCVst) is the variability of the estimated abundance index across the survey replicates and is defined as the standard deviation of the estimated indices of abundance normalized by the true value" (Oyafuso et al 2023)
```{r}
# trueN_stratmu <- mean(trueN_strat$stratmu)
# 
comp_1 |>
 group_by(sim, year) |>
 summarise(cv = sqrt(est_stratvar) / I)

```

Bias
```{r}
# diff <- estN_strat |> 
#   #rename(replicate = "sim") |>
#   group_by(sim, year) |> 
#   summarise(diff = stratmu - trueN_stratmu)
# 
# diff |> 
#   group_by(sim) |>
#   summarise(bias = sum(diff)/length(diff))

```



```{r save data, include = FALSE}

```

