---
title: "09 - Ratio Estimator"
author: "Catalina Roman"
date: "2026-02-11"
output:
  pdf_document:
    latex_engine: pdflatex
    fig_caption: yes
    keep_tex: yes
    
    
header-includes:
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# OBJECTIVE

For a given species, calculate a ratio estimator as a potential mitigation approach to adjust for the loss of sampling coverage caused by offshore wind energy development.

Outputs: abundance indices adjusted with the ratio for each population and simulation of the projection.


# PACKAGES

```{r}
library(tidyverse)
library(here)
library(sdmTMB)
source(here("R", "sim_stratmean_fn.R"))
library(kableExtra)
```


# DATA SET UP

## Directories

```{r}
sseep.analysis <- "C:/Users/croman1/Desktop/UMassD/sseep-analysis"
dist.dat <- here("data", "rds", "dists")
survdat <- here("data", "rds", "survdat")
surv.prods <- here("data", "rds", "surv-prods")
plots <- here("outputs", "plots")
```

## Parameters

```{r}
species <- "scup"
season  <- "fall"
ages      <- 0:7
years     <- 1:15
nsims   <- 1:100
ids     <- sprintf("%03d", nsims)
```

## Data 

```{r load-data, include = FALSE}
#survdat_sq <- map(ids, ~readRDS(here(survdat, sprintf("%s_%s_%s_25_sq_survey.rds", species, season, .x))))
survdat_sq <- map(ids, function(id) {
  x <- readRDS(here(survdat, sprintf("%s_%s_%s_25_sq_survey.rds",
                                     species, season, id)))
  out <- x$setdet #loas only setdet data
  rm(x); gc()
  out
})

survdat_precl <- map(ids, ~readRDS(here(survdat, sprintf("%s_%s_%s_25_precl_survey.rds", species, season, .x)))) 
dist <- map(ids, ~readRDS(here(dist.dat, sprintf("%s_%s_%s_dist-only.rds", species, season, .x))))
```

# Area weights for each strata
```{r}
strata_wts <- readRDS(here(sseep.analysis, "data", "rds", "active_strata_wts.rds")) |>
  rename(strat = STRATUM)
```

# Total survey area
```{r}
survey_area <- as.integer(sum(strata_wts$Area_SqNm))
```

# ABUDNANCE INDEX
calculate the abundance index and relative abundance index for each of the scenarios

## Status Quo Survey

```{r}
#extract the strata that were used to predict spatial distributions in sdmTMB
strat <- map(dist, ~unique(.$strat))

ihat_sq_all <- map2_dfr(survdat_sq, seq_along(survdat_sq), function(surv, pop_num) {
  surv |> #it is already a data.table
    as_tibble() |>
    filter(strat %in% unlist(strat)) |>
    group_by(sim, year, strat) |>
    summarise(towct = length(unique(set)),
              mu = sum(n)/towct,
              var = ifelse(towct <2 , 0, var(n)),
              .groups = "drop") |>
    left_join(strata_wts, by = "strat") |>
    mutate(W = Area_SqNm / survey_area,
           wt_mu = W * mu,
           wt_var = ifelse (towct < 2, 0, (W^2) * var / towct)) |>
    group_by(sim, year) |>
    summarise(stratmu  = sum(wt_mu),
              stratvar = sum(wt_var),
              cv = sqrt(stratvar) / stratmu,
              .groups = "drop") |>
    mutate(scenario = "Status Quo",
           pop = pop_num)
})


# Group by pop (and sim if needed) to compute rel_ihat within each population
ihat_sq_all <- ihat_sq_all %>%
  group_by(pop, sim) %>%
  mutate(rel_ihat = stratmu / mean(stratmu),
         log_se = sqrt(log(1 + cv^2)),
         log_lower = log(stratmu) - 1.96 * log_se,
         log_upper = log(stratmu) + 1.96 * log_se,
         ci_lower = exp(log_lower),
         ci_upper = exp(log_upper)) %>%
  ungroup()

ihat_sq_all |> 
  dplyr::slice_head(n = 20) |>
  knitr::kable(format="latex", booktabs=TRUE, digits=2,
              caption = "Status Quo Abundance Index (first 20 rows)") |>
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down"))
```


## Status quo stratum means (baseline)

```{r}
sq_mu_y <- map2_dfr(survdat_sq, seq_along(survdat_sq), function(surv, pop_num) {
  surv |>
    as_tibble() |>
    filter(strat %in% unlist(strat), year %in% 1:5) |>
    group_by(pop = pop_num, sim, year, strat) |>
    summarise(
      towct = n_distinct(set),
      mu_sq = sum(n) / towct,
      .groups = "drop"
    )
})


sq_mu_y |> 
   dplyr::slice_head(n = 20) |>
  knitr::kable(format="latex", booktabs=TRUE, digits=2,
              caption = "Status Quo mean years 1-5 (first 20 rows)") |>
  kableExtra::kable_styling(latex_options = c("H"), font_size = 8)
```

## Preclusion Survey

```{r}
ihat_precl_all <- map2_dfr(survdat_precl, seq_along(survdat_precl), function(surv, pop_num) {
  surv |> 
    as_tibble() |>
    filter(strat %in% unlist(strat)) |>
    group_by(sim, year, strat) |>
    summarise(towct = length(unique(set)),
              mu = sum(n)/towct,
              var = ifelse(towct <2 , 0, var(n)),
              .groups = "drop") |>
    left_join(strata_wts, by = "strat") |>
    mutate(W = Area_SqNm / survey_area,
           wt_mu = W * mu,
           wt_var = ifelse (towct < 2, 0, (W^2) * var / towct)) |>
    group_by(sim, year) |>
    summarise(stratmu  = sum(wt_mu),
              stratvar = sum(wt_var),
              cv = sqrt(stratvar) / stratmu,
              .groups = "drop") |>
    mutate(scenario = "Preclusion",
           pop = pop_num)
})

# Group by pop (and sim if needed) to compute rel_ihat within each population
ihat_precl_all <- ihat_precl_all %>%
  group_by(pop, sim) %>%
  mutate(rel_ihat = stratmu / mean(stratmu),
         log_se = sqrt(log(1 + cv^2)),
         log_lower = log(stratmu) - 1.96 * log_se,
         log_upper = log(stratmu) + 1.96 * log_se,
         ci_lower = exp(log_lower),
         ci_upper = exp(log_upper)) %>%
  ungroup()

ihat_precl_all |> 
  dplyr::slice_head(n = 20) |>
  knitr::kable(format="latex", booktabs=TRUE, digits=2,
              caption = "Preclusion Abundance Index (first 20 rows)") |>
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down"))
```

## Preclusion stratum means for baseline

```{r}
precl_mu_y <- map2_dfr(survdat_sq, seq_along(survdat_sq), function(surv, pop_num) {
  surv |>
    as_tibble() |>
    filter(strat %in% unlist(strat), year %in% 1:5, AREA_CODE != 1) |>
    group_by(pop = pop_num, sim, year, strat) |>
    summarise(
      towct = n_distinct(set),
      mu_precl = sum(n) / towct,
      .groups = "drop"
    )
})

precl_mu_y |> 
   dplyr::slice_head(n = 20) |>
  knitr::kable(format="latex", booktabs=TRUE, digits=2,
              caption = "Preclusion mean years 1-5 (first 20 rows)") |>
   kableExtra::kable_styling(latex_options = c("H"), font_size = 8)
```



# Ratio calculation (y_tot/y_preclusion)

```{r}
# ratio_tbl <- sq_mu |>
#   left_join(precl_mu, by = c("pop", "sim", "strat"))
# 
# ratio_tbl <- ratio_tbl |>
#   mutate(ratio = ifelse(is.na(mu_precl) | mu_precl == 0, 1, mu_tot / mu_precl))
# 



ratio_tbl <- sq_mu_y |>
  group_by(pop, sim, year) |>
  summarise(mean_sq = mean(mu_sq, na.rm = TRUE), .groups = "drop") |>
  left_join(
    precl_mu_y |>
      group_by(pop, sim, year) |>
      summarise(mean_precl = mean(mu_precl, na.rm = TRUE), .groups = "drop"),
    by = c("pop","sim","year")
  ) |>
  mutate(ratio = ifelse(is.na(mean_precl) | mean_precl == 0, 1, mean_sq / mean_precl))

ratio_tbl |>
  dplyr::slice_head(n = 20) |>
  knitr::kable(format="latex", booktabs=TRUE, digits=2,
              caption = "Ratio (first 20 rows)") |>
  kableExtra::kable_styling(latex_options = c("H"), font_size = 8)

```



# Calculate expanded index

```{r}
ihat_precl_exp_all <- map2_dfr(survdat_precl, seq_along(survdat_precl), function(surv, pop_num) {
    surv |>
    as_tibble() |>
    filter(strat %in% unlist(strat), year %in% 6:15) |>
    group_by(sim, year, strat) |>
    summarise(towct = length(unique(set)),
              mu = sum(n)/towct,
              var = ifelse(towct < 2, 0, var(n)),
             .groups = "drop") |>
    mutate(pop = pop_num) |>
    left_join(ratio_tbl |> select(pop, sim, year, ratio),
              by = c("year","pop","sim")) |>
    mutate(ratio = ifelse(is.na(ratio), 1, ratio),   
      mu_exp = mu * ratio) |>
    left_join(strata_wts, by = "strat") |>
    mutate(W = Area_SqNm / survey_area,
           wt_mu = W * mu_exp,
           wt_var = ifelse (towct < 2, 0, (W^2) * var / towct)) |>
    group_by(sim, year) |>
    summarise(stratmu  = sum(wt_mu),
              stratvar = sum(wt_var),
              cv = sqrt(stratvar) / stratmu,
              .groups = "drop") |>
    mutate( scenario = "Preclusion_expanded",
      pop = pop_num)
})


ihat_precl_exp_all <- ihat_precl_exp_all %>%
  group_by(pop, sim) %>%
  mutate(rel_ihat = stratmu / mean(stratmu),
         log_se = sqrt(log(1 + cv^2)),
         log_lower = log(stratmu) - 1.96 * log_se,
         log_upper = log(stratmu) + 1.96 * log_se,
         ci_lower = exp(log_lower),
         ci_upper = exp(log_upper)) %>%
  ungroup()

ihat_precl_exp_all |>   
  dplyr::slice_head(n = 20) |>
  knitr::kable(format="latex", booktabs=TRUE, digits=2,
              caption = "Preclusion expanded (first 20 rows)") |>
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down"))

```




